Uncertainty Estimation Methods for Countering Attacks on Machine-Generated Text Detectors

### 2025-04-10 paper/main.tex
error analize

\textbf{The first approximation:} Let's start with a simple check of the distribution of the average entropy and perplexity metrics depending on the class label. It's interesting to see if they differ.  To do this, we will use part of the M4GT\citep{wang2024m4gt} dataset as data.

\begin{figure}[bhtp]
	\includegraphics[width=\textwidth]{arxiv.png}
	\caption{Perplexity distribution differs for different models.}
	\label{fig:schema}
\end{figure}

We processed texts from both labels using the LLM (Llama-3.1-8B-Instruct) and utilized contextual logits to compute two metrics: perplexity and entropy. For consistency of the experiment, all data was taken from ARXIV. By analyzing the distribution of points on the plane (Figure 1). It is noticeable that the distribution of models is very different. The texts of different models are clustered according to the values of perplexity, which is a very interesting result. Dolly has similar entropy as humans, while chatGPT, bloomz, davinci and gpt4 is very different. This finding is promising, as it suggests that the distinction between MG and HW text distributions can potentially be detected using uncertainty estimation (UE). In the following sections, we explore this idea further to develop a more robust detection framework.

>>>>

The primary objective is to compare binary classification based on uncertainty estimates with alternative methods for detecting machine-generated text. To do this, we will use part of the M4GT\citep{wang2024m4gt} dataset as data.

\begin{table}[ht]
\centering
\begin{tabular}{llrrrrr}
\toprule
\textbf{Metric} & \textbf{Label} & \textbf{Mean} & \textbf{Median} & \textbf{Std} & \textbf{Min} & \textbf{Max} \\
\midrule
mean entropy   & Human   & 0.0045   & 0.0045   & 0.0005   & 0.0023   & 0.0060   \\
mean entropy   & Machine & 0.0034   & 0.0035   & 0.0009   & 0.0004   & 0.0067   \\
perplexity      & Human   & 2.6851   & 2.6658   & 0.3378   & 1.6291   & 3.8900   \\
perplexity      & Machine & 2.1756   & 2.1370   & 0.4377   & 1.0716   & 4.5481   \\
mc entropy     & Human   & 429.206  & 452.375  & 115.155  & 164.375  & 695.500  \\
mc entropy     & Machine & 216.063  & 190.125  & 98.2335  & 19.0625  & 674.500  \\
mahalanobis     & Human   & 1465.920 & 1400.880 & 350.174  & 357.836  & 4817.470 \\
mahalanobis     & Machine & 2029.660 & 1983.840 & 469.339  & 996.327  & 6088.490 \\
\bottomrule
\end{tabular}
\caption{Metrics for human and machine-generated texts}
\label{tab:metric-by-label}
\end{table}

We processed texts from both labels using the LLM (Llama-3.1-8B-Instruct) and utilized contextual logits to compute four metrics: perplexity, mean token entropy, monte carlo sequence entropy and mahalanobis distance. For consistency of the experiment, all data was taken from Reddit. The results can be seen in Table 1. Some metrics show pronounced variation between human- and machine-generated texts. This finding is promising, as it suggests that the distinction between MG and HW text distributions can potentially be detected using uncertainty estimation (UE). In the following sections, we explore this idea further to develop a more robust detection framework.

The next step involves training a simple neural network classifier. A lightweight architecture comprising several linear layers was selected and implemented. We will compare this model with RoBERTa for sequence classification. The model will be fine-tuned on the same training dataset, and the results will be compared.

\begin{figure}[ht]
\centering
\includegraphics[width=0.5\textwidth]{loss.png}
\caption{Loss for Simple Classifier}
\label{fig:image}
\end{figure}

\begin{table}[ht]
\centering
\begin{tabular}{|l|c|c|c|c|}
\hline
\textbf{Model} & \textbf{Accuracy} & \textbf{Precision} & \textbf{Recall} & \textbf{Train Time (s)} \\
\hline
RobertaForSequenceClassification & 0.9901 & 0.9901 & 0.9992 & 594.4377 \\
neural network classifier with uncertainty & 0.9136 & 0.9455 & 0.9542 & 30.1410 \\
\hline
\end{tabular}
\caption{Comparison of model performance metrics}
\label{tab:model-comparison}
\end{table}

Table 2 presents the results. While our model achieves a good accuracy, it does not outperform RoBERTa. This discrepancy may be attributed to limitations in the classifier itself. Therefore, by selecting a different classifier model, performance could be improved. However, it is noteworthy that the training time of our classifier is significantly shorter.
